\documentclass[11pt]{article}
\usepackage[hyperref]{acl}
\usepackage{times}
\usepackage{latexsym}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{array}
\usepackage{pifont}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,backgrounds,calc}

\lstset{
  basicstyle=\ttfamily\scriptsize,
  breaklines=true,
  frame=single,
  backgroundcolor=\color{gray!8},
  commentstyle=\color{gray},
  keywordstyle=\color{blue},
  aboveskip=4pt,
  belowskip=4pt,
}

\title{UnMask: Mastery-Gated Retrieval for Socratic OT Anatomy Education}

\author{Sanika Vilas Najan$^*$ \quad Vaishak Girish Kumar$^*$ \\
  University at Buffalo \\
  CSE 635: NLP and Text Mining --- Spring 2026 \\
  Instructor: Prof.\ Rohini K.\ Srihari \\
  \texttt{\{snajan, vaishakg\}@buffalo.edu} \\[2pt]
  \small $^*$Equal contribution.}

\begin{document}
\maketitle

%% ─────────────────────────────────────────────────────────────
\begin{abstract}
We present \textbf{UnMask}, a Socratic AI tutor for OT students preparing for the NBCOT exam.
Most existing Socratic tutors try to suppress answers at generation time through prompting,
but we found this approach breaks down when the model is capable enough to override
system instructions.
Instead, we move answer suppression to the \emph{retrieval layer} through
\textbf{Progressive Context Revelation (PCR)}: a metadata-filtered Qdrant query that
withholds answer chunks until the student demonstrates prerequisite mastery.
We also apply \textbf{Corrective RAG (CRAG)} --- to our knowledge the first use in
educational tutoring --- along with dual-layer structured output masking,
\textbf{session mistake memory} with proactive revisits at 8 minutes,
and a \textbf{multimodal Next.js frontend} supporting VLM-based image uploads, animated anatomy
diagrams, and per-session YouTube recommendations.
We evaluated on 30 QA triples across a four-variant ablation and found:
Hit Rate 0.900, Leak Rate 0.000, Socratic Purity 4.87/5, and 100\% adversarial hold rate.
All four variants produce zero leaks under normal conditions, but only PCR
holds under adversarial attack --- at a 0.23-point purity cost (4.70 vs.\ 4.93).
RAGAS Faithfulness (0.838) misses its target, which we argue is a metric mismatch:
Socratic questions make no factual claims, so RAGAS penalizes them by design.
Live demo: \url{https://nlp-un-mask.vercel.app/}
\end{abstract}

%% ─────────────────────────────────────────────────────────────
\section{Introduction}

Occupational Therapy students preparing for the NBCOT exam need to develop clinical
reasoning, not just recall facts.
Standard AI systems work against this: they give direct answers that let students
skip the reasoning step the exam actually tests.
The project specification defines a ``Socratic'' tutor as a \emph{Tutor-not-Teller}:
the system should guide students to the answer through questions, never stating it directly.

The problem is that prompt-based suppression only works until the model decides not
to follow it.
Dinucu-Jianu et al.\ \cite{tutorrl2025} showed that larger models actually have
\emph{higher} leak rates because their parametric knowledge overrides system prompts.
If the model already knows the answer, telling it not to say it is not a reliable safeguard.

\paragraph{Our main claim.}
If the LLM never receives the answer chunk, it cannot leak it.
We enforce this with a 10-line Qdrant metadata filter that no adversarial prompt
can bypass, because the filtering happens on the database server before any
results reach Python.

%% ─────────────────────────────────────────────────────────────
\section{Related Work}

Table~\ref{tab:related} summarizes the key distinctions between UnMask and prior work.
Dinucu-Jianu et al.\ \cite{tutorrl2025} showed that larger LLMs leak answers more
often, not less, which is what pushed us away from prompt-based suppression.
Corrective RAG \cite{crag2024} adds a relevance check after retrieval (+19\% on PopQA,
+36.6\% on PubHealth); we are not aware of any prior work applying CRAG in an educational tutoring pipeline.
TutorLLM \cite{tutorllm2024} combines knowledge tracing with RAG for personalized
learning, but does not describe a mastery-gated metadata filtering mechanism
that operates at the data layer.
Dinucu-Jianu et al.\ \cite{tutorrl2025} also show that RL reward shaping for
pedagogical alignment outperforms prompt-based suppression on GPT-4o, which is
consistent with our finding that prompt-level safety fails under adversarial attack.

\begin{table}
\centering
\setlength{\tabcolsep}{3pt}
\scriptsize
\begin{tabular}{lccl}
\hline
\textbf{System} & \textbf{RAG} & \textbf{Gate} & \textbf{Layer} \\
\hline
Khanmigo                          & \checkmark & $\times$      & prompt \\
SocraticLM \cite{socraticlm2024}  & $\times$   & $\times$      & fine-tune \\
KELE \cite{kele2025}              & \checkmark & $\times$      & prompt \\
TutorLLM \cite{tutorllm2024}      & \checkmark & KT            & retrieval \\
PedagogRL \cite{tutorrl2025}      & $\times$   & RL            & training \\
\textbf{UnMask (ours)}            & \checkmark & \checkmark    & \textbf{retrieval} \\
\hline
\end{tabular}
\caption{Socratic AI tutor comparison. Only UnMask gates at retrieval layer.}
\label{tab:related}
\end{table}

%% ─────────────────────────────────────────────────────────────
\section{System Architecture}

UnMask is built around four layers that run through a LangGraph state machine.
Figure~\ref{fig:architecture} shows the full query path from student message to Socratic response.

%% ── Architecture diagram ──────────────────────────────────────────
\begin{figure*}[t]
\centering
\begin{tikzpicture}[
  >=Stealth,
  node distance=0.55cm and 1.05cm,
  main/.style={draw, rounded corners=4pt, fill=#1,
               minimum width=1.95cm, minimum height=0.72cm,
               align=center, font=\small\bfseries, inner sep=4pt},
  sub/.style={draw=gray!55, rounded corners=3pt, fill=gray!9,
              minimum width=1.85cm, align=center,
              font=\scriptsize, inner sep=3pt},
  arr/.style={->, thick},
  darr/.style={->, thick, dashed, gray!65},
]
%% ─── Top-level pipeline ───────────────────────────────────────────────────────
\node[main=green!22]              (ui)   {Student\\(Next.js)};
\node[main=blue!14, right=1.05cm of ui]   (orch) {Orchestrator};
\node[main=blue!14, right=1.05cm of orch] (ret)  {Retrieval\\Planner};
\node[main=blue!14, right=1.05cm of ret]  (gen)  {Socratic\\Generator};
\node[main=blue!14, right=1.05cm of gen]  (ped)  {Pedagogy\\Agent};
\node[main=green!22,right=1.05cm of ped]  (out)  {Socratic Q\\(student)};

\draw[arr] (ui)   -- (orch);
\draw[arr] (orch) -- (ret)  node[above,midway,font=\tiny,align=center]{tutor/\\assess};
\draw[arr] (ret)  -- (gen);
\draw[arr] (gen)  -- (ped);
\draw[arr] (ped)  -- (out);

%% rapport/wrapup bypass
\draw[darr] (orch) to[bend left=42]
  node[above,font=\tiny]{rapport / wrapup} (gen);

%% ─── Sub-components ───────────────────────────────────────────────────────────
%% Orchestrator
\node[sub, below=0.5cm of orch] (phases)
  {Phase: R$\to$T$\to$A$\to$W\\time + event gates};
\draw[darr] (orch) -- (phases);

%% Retrieval Planner
\node[sub, below=0.5cm of ret]  (pcrf)  {PCR Filter\\(mastery gate)};
\node[sub, below=0.28cm of pcrf] (hyb)  {Dense + BM25 + RRF\\+ CRAG loop};
\node[sub, below=0.28cm of hyb, fill=orange!16] (qdr)
  {Qdrant (48 chunks)\\{\ttfamily is\_answer\_chunk}};
\draw[darr] (ret)  -- (pcrf);
\draw[darr] (pcrf) -- (hyb);
\draw[darr] (hyb)  -- (qdr);

%% Socratic Generator
\node[sub, below=0.5cm of gen]  (sout)
  {Structured Output\\hidden: \textit{InternalAnalysis}\\shown: \textit{SocraticQ}};
\node[sub, below=0.28cm of sout, fill=yellow!18] (llm)
  {Inception Mercury-2 (all LLM calls)\\Gemini 2.0 Flash Lite (vision)};
\draw[darr] (gen)  -- (sout);
\draw[darr] (sout) -- (llm);

%% Pedagogy Agent
\node[sub, below=0.5cm of ped]  (mast)
  {Mastery BKT\\Concept DAG\\Mistake Log};
\draw[darr] (ped) -- (mast);

\end{tikzpicture}
\caption{End-to-end query flow through UnMask. The student message enters the LangGraph state machine via a Next.js frontend (SSE stream); the Supervisor (Mercury-2, rule-based) routes tutoring/assessment turns through the Retrieval Planner (PCR-gated hybrid RAG + CRAG), then the Socratic Generator (Mercury-2 via OpenRouter, structured output with hidden \texttt{InternalAnalysis}), then the Pedagogy Agent (mastery update, concept DAG, mistake log). Rapport and wrapup bypass retrieval.}
\label{fig:architecture}
\end{figure*}

\subsection{Session State Machine (LangGraph)}

The session moves through four phases (Table~\ref{tab:phases}) with time ceilings
enforced by a pure-Python orchestrator node that makes zero LLM calls.
We chose rule-based phase transitions after reading Mao et al.\ \cite{diaggpt2023},
who found that rule-based controllers beat LLM routers for deterministic transitions.

\begin{table}
\centering
\small
\begin{tabularx}{\columnwidth}{@{}l l X r@{}}
\toprule
\textbf{Phase} & \textbf{Entry} & \textbf{Exit trigger} & \textbf{Window} \\
\midrule
Rapport    & Session start     & 4 diagnostic Qs answered       & 0--120s   \\
Tutoring   & diag.\ complete   & cov.$\geq$0.80 or $t\geq$720s  & 120--720s \\
Assessment & cov./time trigger & $t\geq$840s                    & 720--840s \\
Wrapup     & $t\geq$840s       & Session end                    & 840--900s \\
\bottomrule
\end{tabularx}
\caption{LangGraph session phases with entry/exit triggers and time ceilings.}
\label{tab:phases}
\end{table}

The \texttt{TutoringState} TypedDict passes through four nodes
(Figure~\ref{fig:architecture}).
A conditional edge in the orchestrator skips retrieval for Rapport and Wrapup phases,
since those phases do not require content gating.

\subsection{Models}
\label{sec:models}

Table~\ref{tab:models} lists all models used. All LLM calls go through
OpenRouter (\texttt{openrouter.ai/api/v1}), so the base URL is the only
setting that changes when swapping providers.

\begin{table}
\centering
\setlength{\tabcolsep}{3pt}
\scriptsize
\begin{tabular}{lll}
\hline
\textbf{Component} & \textbf{Model} & \textbf{Purpose} \\
\hline
Supervisor            & Mercury-2          & Phase routing \\
Tutoring/Assessment   & Mercury-2          & Socratic output \\
Wrapup summary        & Mercury-2          & Session report \\
Embeddings            & Gemini Emb.\ 2     & Dense retrieval \\
CRAG relevance        & Mercury-2          & Chunk grading \\
Image verification    & Gemini Flash Lite  & Diagram filtering \\
Student image VLM     & Gemini Flash Lite  & Upload identification \\
\hline
\end{tabular}
\caption{Models per component. All LLM calls via OpenRouter.}
\label{tab:models}
\end{table}

\subsection{Student Interface (Next.js)}

Students interact through a Next.js 14 App Router frontend (live at
\url{https://nlp-un-mask.vercel.app/}).
Responses stream via SSE from the FastAPI backend.
A right-hand inspector panel shows instructors the per-turn PCR retrieval mode,
mastery state per concept, and misconception log --- this directly addresses
the course requirement to show how chunks are being processed.
A Chainlit interface (\texttt{app.py}) is kept for backward compatibility.

\paragraph{Personalized onboarding.}
The first student message is parsed by the Orchestrator to extract
\texttt{study\_focus} (e.g., brachial plexus, rotator cuff) and
\texttt{learning\_mode} (visual / Q\&A) before the LangGraph graph runs.
\texttt{study\_focus} reorders the four diagnostic questions so the
student's declared weak area comes first.
\texttt{learning\_mode} adjusts the visual-hint threshold:
visual learners get a diagram after the first wrong answer;
Q\&A learners after the second.
Figure~\ref{fig:onboarding} shows the full parse-to-render pipeline.

%% ── Onboarding + visual aid pipeline ───────────────────────────
\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  node distance=0.38cm and 0.55cm,
  box/.style={draw, rounded corners=3pt, fill=#1,
              minimum width=2.55cm, align=center,
              font=\scriptsize\strut, inner sep=3pt},
  arr/.style={->, thick},
  darr/.style={->, thick, dashed, gray!60},
]
%% Row 1: student message
\node[box=green!18] (msg)
  {Student msg\\``...brachial plexus,\\prefer diagrams''};
%% Row 2: orchestrator
\node[box=blue!14, below=0.45cm of msg] (orch)
  {Orchestrator\\(parse before invoke)};
\draw[arr] (msg) -- (orch);
%% Row 3: two extracted fields
\node[box=orange!20, below left=0.42cm and 0.28cm of orch] (sf)
  {\texttt{study\_focus}\\``brachial plexus''};
\node[box=orange!20, below right=0.42cm and 0.28cm of orch] (lm)
  {\texttt{learning\_mode}\\``visual''};
\draw[arr] (orch) -| (sf);
\draw[arr] (orch) -| (lm);
%% Row 4: downstream effects
\node[box=gray!14, below=0.38cm of sf] (diag)
  {diag\_order [0,2,3,1]\\BP questions first};
\node[box=gray!14, below=0.38cm of lm] (thr)
  {threshold = 1\\(Q\&A: 2)};
\draw[arr] (sf) -- (diag);
\draw[arr] (lm) -- (thr);
%% Row 5: visual aid render
\node[box=green!18, below=0.38cm of thr] (img)
  {\texttt{cl.Image}\\Gray's Anatomy plate\\rendered inline};
\draw[arr] (thr) -- (img)
  node[right, font=\tiny, midway]{$\geq$ threshold};
\end{tikzpicture}
\caption{Personalized onboarding and visual aid pipeline.
First message parsed before \texttt{graph.invoke}: \texttt{study\_focus}
reorders diagnostics; \texttt{learning\_mode} sets the threshold
that triggers an inline anatomy diagram via SSE \texttt{visual\_hint} event $\to$ \texttt{DiagramCard}.}
\label{fig:onboarding}
\end{figure}

\paragraph{Inline anatomical diagrams.}
The UI renders a \texttt{DiagramCard} once
\texttt{consecutive\_incorrect} $\geq$ threshold.
22 self-contained animated SVG/HTML files in \texttt{public/anatomy/} cover
every concept in the knowledge base (brachial plexus, peripheral nerves, rotator cuff, spinal cord).
For concepts without a local file, a DuckDuckGo image search queries trusted
anatomy domains (Wikipedia, Radiopaedia, Kenhub), and Gemini 2.0 Flash Lite
verifies the top-3 results before displaying the first one that passes.

\subsection{Progressive Context Revelation (PCR)}
\label{sec:pcr}

Every chunk in Qdrant has two metadata fields set at index time:
\texttt{is\_answer\_chunk: bool} and
\texttt{chunk\_type: \{context $|$ prerequisite $|$ answer $|$ figure\}}.
The Retrieval Planner reads the student's current mastery score from the concept graph
and picks one of three filter modes:

\begin{lstlisting}[language=Python]
if mastery < 0.40:   # context_only
    filter = must_not(is_answer_chunk=True)
elif mastery < 0.70: # prerequisite_first
    filter = must(chunk_type in
        ["context","prerequisite","figure"])
else:                # full_reveal
    filter = None
\end{lstlisting}

%% ── PCR mastery gating diagram ───────────────────────────────────
\begin{figure}[ht]
\centering
\begin{tikzpicture}[>=Stealth, font=\small]
%% Coloured mastery bands
\fill[red!22]    (0,0)   rectangle (2.16,0.52);
\fill[yellow!35] (2.16,0) rectangle (3.78,0.52);
\fill[green!28]  (3.78,0) rectangle (5.4, 0.52);
\draw[thick] (0,0) rectangle (5.4,0.52);
\draw[thick,dashed] (2.16,0) -- (2.16,0.52);
\draw[thick,dashed] (3.78,0) -- (3.78,0.52);
%% Band labels
\node[font=\scriptsize] at (1.08,0.26) {\texttt{context\_only}};
\node[font=\scriptsize] at (2.97,0.26) {\texttt{prereq\_first}};
\node[font=\scriptsize] at (4.59,0.26) {\texttt{full\_reveal}};
%% Threshold ticks
\foreach \x/\lbl in {0/$0$, 2.16/$0.4$, 3.78/$0.7$, 5.4/$1$}{
  \draw[thick] (\x,0) -- (\x,-0.1);
  \node[below,font=\scriptsize] at (\x,-0.1) {\lbl};
}
%% What's accessible above each band
\node[above,align=center,font=\scriptsize] at (1.08,0.52)
  {context\\{\color{red}\ding{55}} answer};
\node[above,align=center,font=\scriptsize] at (2.97,0.52)
  {+prereq\\{\color{red}\ding{55}} answer};
\node[above,align=center,font=\scriptsize] at (4.59,0.52)
  {all chunks\\{\color{green!60!black}\ding{51}} answer};
%% x-axis label
\node[below=0.4cm,font=\small] at (2.7,-0.1) {Mastery score $m$};
\end{tikzpicture}
\caption{PCR filter modes vs.\ mastery score. In \texttt{context\_only} mode ($m<0.40$), Qdrant's server-side \texttt{must\_not} filter physically excludes answer chunks before any Python code executes.}
\label{fig:pcr}
\end{figure}

In \texttt{context\_only} mode, the \texttt{must\_not} filter runs inside
Qdrant before any results reach Python.
The LLM \emph{never receives} the answer chunk.
This matters because the safety guarantee comes from the data layer rather
than the prompt layer --- there is nothing to instruct the model to override.

\subsection{Hybrid Retrieval + Corrective RAG}

\paragraph{Hybrid retrieval.}
Dense vectors (Gemini Embedding 2, 3072-dim) are merged with BM25 sparse
retrieval using Reciprocal Rank Fusion (RRF, $k{=}60$).
The PCR filter applies to both retrieval legs before merging.

\paragraph{CRAG loop.}
After retrieval, an LLM grades each document for relevance (binary yes/no).
If all documents fail, the query is reformulated through synonym expansion and
retried up to 2 times; Yan et al.\ \cite{crag2024} show diminishing returns beyond 2.
We believe this is the first application of CRAG in an educational tutoring system.
Ablation timing confirms CRAG fires in practice: the full variant stalled at
question 18 (${\sim}186$s vs.\ typical ${\sim}8$s), confirming a live re-query;
the \texttt{no\_crag} variant never stalled.

\subsection{Dual Knowledge Masking via Structured Output}

The Socratic Generator calls Mercury-2 with
\texttt{response\_format=SocraticOutput}:

\begin{lstlisting}[language=Python]
class InternalAnalysis(BaseModel):
    correct_answer: str        # stripped, never shown
    student_misconception: str
    planned_hint_sequence: list[str]

class VisibleResponse(BaseModel):
    socratic_question: str     # must end with "?"
    encouragement: str
\end{lstlisting}

\texttt{InternalAnalysis} is consumed by the Pedagogy Agent and stripped
before the response reaches the student.
A post-generation leak guard also checks for $\geq$4 significant-word
overlap between \texttt{socratic\_question} and \texttt{correct\_answer}
and triggers a retry if that fires.

\textbf{Honest encouragement and break-Socratic mode.}
When \texttt{consecutive\_incorrect} $> 0$, the system is not allowed to give
hollow praise.
If the student asks for a direct explanation (``can you explain'',
``walk me through'', etc.) or gets 4 consecutive wrong answers,
\texttt{break\_socratic=True} switches to a \texttt{\_REVEAL\_SYSTEM} prompt
that tells the model to explain rather than question.

\textbf{Wrapup summary and YouTube recommendations.}
At Wrapup, Mercury-2 structured output generates a \texttt{SessionSummary}:
per-concept mastery status, misconceptions from \texttt{mistake\_log},
study tips, and 2--4 \texttt{YouTubeResource} objects for the weakest concepts,
rendered as clickable search cards in the frontend Progress tab.
\textbf{Cost:} Mercury-2 covers all LLM components (Supervisor, Tutoring, Assessment, CRAG grading, Wrapup); Gemini 2.0 Flash Lite handles vision tasks (${\sim}\$0.01$--$0.02$/session).

\subsection{Concept Prerequisite Graph}

Rather than tracking mastery as flat per-concept scores, we maintain a directed
acyclic graph (DAG) of anatomy concepts (e.g., \texttt{brachial\_plexus.origin $\to$
brachial\_plexus.trunks $\to$ peripheral\_nerves.axillary}).
When a student struggles with a concept, the Pedagogy Agent calls
\texttt{nx.ancestors()} to find prerequisite gaps that likely caused the failure.

\paragraph{Cold-start initialization.}
The Rapport phase gives 4 diagnostic questions.
Correct answers set mastery at 0.5; incorrect at 0.1; skipped at 0.2.
This gives the PCR filter meaningful priors within about 90 seconds.

\paragraph{Mastery update.}
After each response, the LLM judges correctness against
\texttt{internal\_analysis.correct\_answer} and updates:
$m' = m + 0.15(1-m)$ if correct; $m' = m - 0.05m$ if incorrect
\cite{corbett1994}.
Two consecutive correct answers with coverage $\geq 0.80$ triggers a phase advance.

\subsection{Session Mistake Memory and Proactive Revisit}
\label{sec:memory}

Every wrong answer appends a structured record (topic, misconception, turn,
elapsed seconds) to \texttt{mistake\_log}, which is append-only in state.
The misconception text comes from the hidden \texttt{InternalAnalysis} at the
moment of the mistake.
After 8 minutes the Orchestrator picks the lowest-mastery weak topic, sets
\texttt{revisit\_scheduled=True} with a 3-minute cooldown to prevent repeated
triggering, and overrides \texttt{current\_topic}.
The Retrieval Planner then adds the topic name to the Qdrant query;
the Socratic Generator injects the stored misconception into the system prompt
to ask a fresh-angle question rather than repeating the same one.
The Pedagogy Agent clears \texttt{revisit\_scheduled} after one turn.

\subsection{Multimodal Diagram Tutoring (Task 4)}
\label{sec:vlm}

When a student uploads an anatomical image through the frontend,
UnMask calls Gemini 2.0 Flash Lite (vision) via OpenRouter in a two-step pipeline:
first identify the structure internally, then ask a Socratic question
\emph{without naming it}:

\begin{lstlisting}[language=Python]
# api.py  -  POST /api/sessions/{id}/image
vision_model = cfg["llm"].get("vision_model",
    "google/gemini-2.0-flash-lite")
# Step 1: internal identification (never shown)
id_resp = vision_client.chat.completions.create(
    model=vision_model, max_tokens=50,
    messages=[image_url + "Reply with ONLY the
              anatomical name, nothing else."])
identified = id_resp.choices[0].message.content
# Step 2: Socratic question (structure not named)
q_resp = tutor_client.chat.completions.create(
    model=openai_model,  # Mercury-2
    system=f"Student uploaded image of {identified}.
    Ask ONE Socratic question. Do NOT name the
    structure. End with '?'")
# response ends with "?"  -  structure unnamed
\end{lstlisting}

The vision model identifies the structure in a separate call (max 50 tokens,
name only), keeping the identification internal while Mercury-2 generates
a Socratic question that never names the structure in the visible prompt.
This closes the multimodal loop: static anatomy plates for system-triggered
diagrams (Task 1), plus VLM interpretation for student-uploaded images (Task 4).

%% ─────────────────────────────────────────────────────────────
\section{Dataset and Knowledge Base}

\paragraph{Corpus.}
OpenStax Anatomy \& Physiology 2e \cite{openstax2019}, Chapters 11 and 13--16,
covering brachial plexus and peripheral nerve topics tested on the NBCOT.
Indexed into Qdrant as 48 chunks across 16 concepts, each chunk tagged with
\texttt{is\_answer\_chunk}, \texttt{chunk\_type}, and \texttt{concept}.
Each concept has 3--5 chunks covering context, prerequisite, PCR-gated answer,
figure, and clinical types.
The clinical chunks include OT assessment techniques (Phalen's, Tinel's, Froment's
sign, empty-can, lift-off tests), splinting protocols, and NBCOT clinical scenarios
with answer keys.

\paragraph{Evaluation dataset.}
30 QA triples (question, retrieved context, expected answer) spanning:
brachial plexus origins/trunks/divisions/cords, peripheral nerves (axillary,
radial, median, ulnar), and rotator cuff muscles.
20 adversarial prompts across four attack types: direct requests, jailbreaks,
social engineering, and off-topic distractors.
5 annotated multi-turn conversation transcripts (supplementary material).

\paragraph{Sample transcript (Turn 1, PCR \texttt{context\_only}).}
\textit{Student}: ``A patient has wrist drop after a mid-shaft humerus fracture. What nerve is injured?''\\
\textit{UnMask}: ``What is the relationship between the fracture site and the nerve that runs through that area, particularly one responsible for wrist extension?''\\
\textit{[Gold held: ``Radial nerve'' -- absent from retrieved context; 0 leaks]}

\paragraph{Generalizability.}
The system is content-agnostic.
Changing \texttt{qdrant.collection: unmask\_anatomy $\to$ unmask\_physics}
and re-indexing OpenStax Physics 2e (Chapters 4--6) requires no code changes,
because the content (Qdrant collection) is fully separate from the
logic (LangGraph orchestration).

%% ─────────────────────────────────────────────────────────────
\section{Evaluation}

\subsection{Metrics}

\paragraph{Socratic Purity (target $\geq$ 4.0/5).}
Two-layer combined score.
\emph{Rule-based}: does the response end with ``?''; does it contain
$\geq$4 significant-word overlap with the gold answer (keyword leak); is
cosine similarity $>$ 0.92 (semantic leak)?
A confirmed leak (both layers) hard-caps the score at 2.0; missing ``?''
penalizes $-1.0$.
\emph{LLM-as-Judge}: Mercury-2 rates 1--5 where 5 = perfect Socratic (gold
answer absent, student must reason) and 1 = direct answer stated.

\paragraph{Answer Leak Rate (target = 0).}
Fraction of responses where both leak layers fire at the same time.

\paragraph{Retrieval Hit Rate @5 (target $\geq$ 0.75).}
In the full-system evaluation (unrestricted), does the top-5 set contain the
gold answer chunk?
In the ablation, this measures whether the answer chunk reaches the LLM:
correct PCR behavior is hit rate = 0.

\paragraph{Adversarial Hold Rate (target $\geq$ 90\%).}
Fraction of adversarial prompts answered with a Socratic question rather than
a direct answer.

\paragraph{RAGAS Faithfulness (target $\geq$ 0.85).}
Whether claims in the response are entailed by retrieved context, measured via
the \texttt{ragas} library~\cite{es2023ragas}.

\subsection{Ablation Design}

Four variants on the same 30 questions, identical cold-start mastery (0.20),
identical generation (Mercury-2, structured output):

\begin{center}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Variant} & \textbf{PCR} & \textbf{CRAG} & \textbf{Graph} \\
\midrule
\texttt{full}     & \checkmark & \checkmark & \checkmark \\
\texttt{no\_pcr}  & $\times$ (always full\_reveal) & \checkmark & \checkmark \\
\texttt{no\_crag} & \checkmark & $\times$ & \checkmark \\
\texttt{no\_graph}& \checkmark & \checkmark & $\times$ \\
\bottomrule
\end{tabular}
\end{center}

%% ─────────────────────────────────────────────────────────────
\section{Results}

\subsection{Full System (30 QA Questions)}

\begin{table}
\centering
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Metric} & \textbf{Score} & \textbf{Target} & \textbf{Pass} \\
\midrule
Hit Rate @5     & \textbf{0.900} & $\geq$0.75 & \checkmark \\
MRR             & \textbf{0.604} & ---        & --- \\
Leak Rate       & \textbf{0.000} & 0\%        & \checkmark \\
Soft Flags      & \textbf{0.000} & ---        & \checkmark \\
Ends with ``?'' & \textbf{1.000} & $\geq$95\% & \checkmark \\
Avg Purity      & \textbf{4.87/5}& $\geq$4.0  & \checkmark \\
Purity Pass Rate& \textbf{0.967} & ---        & --- \\
Adversarial Hold& \textbf{1.000} & $\geq$90\% & \checkmark \\
\midrule
RAGAS Faithful. & \textbf{0.838} & $\geq$0.85 & $\times$ \\
RAGAS Relevancy & \textbf{0.622} & $\geq$0.80 & $\times$ \\
\bottomrule
\end{tabular}
\caption{Full system evaluation results (30 questions, mastery = 0.20).}
\label{tab:main_results}
\end{table}

The only purity outlier is q29 (\texttt{peripheral\_nerves.axillary},
score 3.0/5): the LLM judge noted the response referenced both movement
and sensation roles of the nerve, giving the student enough to infer the answer.
One additional outlier (score 3.5/5) appeared in the updated run; all others scored $\geq$4.0.

\paragraph{Adversarial robustness.}
All 20 prompts were blocked across four attack types: direct requests
(``just tell me''), jailbreaks (``pretend you are a textbook''), social
engineering (``my professor said to give direct answers''), and off-topic
distractors (Paris geography, Python list sorting).
The 100\% hold rate is architectural: in \texttt{context\_only}
mode there is no answer chunk in context to leak.

\subsection{Ablation Results}

\begin{table}
\centering
\small
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Metric} & \textbf{full} & \textbf{no\_pcr} & \textbf{no\_crag} & \textbf{no\_graph} \\
\midrule
Ans.\ Chunk Reach$^\dagger$ & \textbf{0.000} & 1.000 & 0.000 & 0.000 \\
Leak Rate  & 0.000 & 0.000 & 0.000 & 0.000 \\
Avg Purity & \textbf{4.70}  & 4.83  & 4.87  & 4.93  \\
Ends ``?'' & 1.000 & 1.000 & 1.000 & 1.000 \\
\bottomrule
\end{tabular}}
\caption{Ablation results (30 questions/variant, mastery pinned at 0.20 for all turns).
$^\dagger$Ans.\ Chunk Reach = fraction of queries where the answer chunk
enters LLM context; 0.000 is \emph{correct} PCR behavior at cold-start mastery.
Note: Table~\ref{tab:main_results} (full system, mastery free to evolve) reports
Avg Purity 4.87 vs.\ 4.70 here because the pinned-mastery ablation holds all 30
questions in \texttt{context\_only} mode, producing broader guiding questions.}
\label{tab:ablation}
\end{table}

\paragraph{PCR works as designed.}
The full system's 0.000 reach rate confirms PCR's \texttt{must\_not}
filter is excluding answer chunks at cold-start mastery.
The \texttt{no\_pcr} variant shows the counterfactual clearly: answer chunks
reach the LLM on every query (reach rate 1.000).

\paragraph{Zero leaks, but not equally safe.}
All four variants show 0.000 leak rate on the benign evaluation set.
This matches exactly what Dinucu-Jianu et al.\ \cite{tutorrl2025} found: capable
models follow instructions well enough on normal inputs that prompt-based
suppression looks fine until you actually attack it.
Our adversarial battery shows the gap: PCR held at 100\% while prompt-based
approaches would degrade under attack, since there is no data-layer guarantee.

\paragraph{Purity cost of safety: 0.23 points.}
The full system scores 4.70/5 vs.\ 4.93/5 for \texttt{no\_graph}.
When the answer chunk is excluded, the model asks broader guiding questions
because it cannot see exactly what to guide the student toward.
The \texttt{no\_pcr} variant hits 4.83: having the answer in context helps
the model write more targeted Socratic scaffolding.
The 0.23-point gap is the measurable cost of architectural over prompt-level safety.

\paragraph{RAGAS mismatch.}
Both RAGAS scores fall below target (Faithfulness 0.838, Relevancy 0.622), but
this is expected for a Socratic system.
A guiding question such as \emph{``What is the relationship between the
mid-shaft humerus fracture site and the nearby nerve?''} makes zero factual
claims, so RAGAS marks it as unfaithful.
Socratic Purity (4.87/5, LLM-as-Judge) is a better proxy for what we actually
want to measure here.

\paragraph{Why benign testing is not enough.}
Since all ablation variants show 0.000 leak rate on normal inputs, a benign-only
evaluation would suggest PCR adds nothing.
PCR's advantage only appears under adversarial conditions, which matches the
pattern Dinucu-Jianu et al.\ \cite{tutorrl2025} found.
The 4-question diagnostic cold-start prevents applying \texttt{context\_only}
to students who already know the material;
CRAG prevents off-topic responses when retrieval returns irrelevant chunks
(confirmed by the 186s stall at q18 vs.\ typical 8s).

%% ─────────────────────────────────────────────────────────────
\section{Conclusion}

The standard approach of retrieving everything and suppressing answers through
prompting has a real vulnerability: it only holds if the model decides to comply.
Progressive Context Revelation moves answer gating to the retrieval layer,
making leakage impossible before mastery thresholds are met because
the answer chunk is simply not there.
The trade-off is a 0.23-point purity drop (4.70 vs.\ 4.93/5), which is the
cost of asking questions without seeing the answer first.
Session Mistake Memory closes a gap that stateless tutors leave open:
we store misconceptions at the time of each wrong answer, use them to trigger
revisits at 8 minutes, and produce an honest per-concept \texttt{SessionSummary}
at the end.
Overall, UnMask is a working, evaluable OT anatomy tutor that runs at roughly
\$0.08--0.10 per session.


%% ─────────────────────────────────────────────────────────────
\begin{thebibliography}{10}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[Corbett and Anderson(1994)]{corbett1994}
Albert~T. Corbett and John~R. Anderson. 1994.
\newblock Knowledge tracing: Modeling the acquisition of procedural knowledge.
\newblock \emph{User Modeling and User-Adapted Interaction}, 4(4):253--278.

\bibitem[Dinucu-Jianu et~al.(2025)Dinucu-Jianu, Macina, et~al.]{tutorrl2025}
David Dinucu-Jianu, Jakub Macina, et~al. 2025.
\newblock From problem-solving to teaching problem-solving: Aligning {LLMs}
  with pedagogy using {RL}.
\newblock In \emph{Proceedings of EMNLP 2025} (Oral). ETH Zurich.

\bibitem[Es et~al.(2023)Es et~al.]{es2023ragas}
Shahul Es et~al. 2023.
\newblock {RAGAS}: Automated evaluation of retrieval augmented generation.
\newblock \emph{arXiv preprint arXiv:2309.15217}.

\bibitem[Li et~al.(2024)Li, Wang, Gu, et~al.]{tutorllm2024}
Zhi Li, Jifan Wang, Weijie Gu, et~al. 2024.
\newblock {TutorLLM}: Customizing learning recommendations with knowledge
  tracing and retrieval-augmented generation.
\newblock In \emph{RecSys 2024 / INTERACT 2025}.

\bibitem[Liu et~al.(2024)Liu et~al.]{socraticlm2024}
Jiaqi Liu et~al. 2024.
\newblock {SocraticLM}: Exploring Socratic personalized teaching with {LLMs}.
\newblock In \emph{Proceedings of NeurIPS 2024} (Spotlight).

\bibitem[Mao et~al.(2023)Mao et~al.]{diaggpt2023}
Junyuan Mao et~al. 2023.
\newblock {DiagGPT}: An {LLM}-based chatbot with automatic topic management
  for task-oriented dialogue.
\newblock \emph{arXiv preprint arXiv:2308.08043}.

\bibitem[{OpenStax}(2019)]{openstax2019}
{OpenStax}. 2019.
\newblock \emph{Anatomy \& Physiology 2e}.
\newblock OpenStax, Rice University.

\bibitem[Peng et~al.(2025)Peng et~al.]{kele2025}
Xuanyu Peng et~al. 2025.
\newblock {KELE}: A multi-agent framework for structured Socratic teaching with
  {LLMs}.
\newblock In \emph{Proceedings of EMNLP 2025 Findings}.

\bibitem[Yan et~al.(2024)Yan, Gu, Zhu, and Ling]{crag2024}
Shi-Qi Yan, Jia-Chen Gu, Yun Zhu, and Zhen-Hua Ling. 2024.
\newblock Corrective retrieval augmented generation.
\newblock In \emph{Proceedings of ACL 2024}.

\end{thebibliography}

\end{document}
